\documentclass{article}

\usepackage{fullpage}
\usepackage{syntax}
\usepackage{listings}
\usepackage{listings-golang}
\usepackage{color}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ % add your own preferences
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=t,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=top,frame=bottom,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  otherkeywords={*,...},           % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  title=\lstname,                   % show the filename of files included with \lstinputlisting; also try caption instead of title
  literate={+}{{\texttt{+}}}1 {-}{-}1,
  language=Golang % this is it !
}

\title{Proposal: Go--}
\author{Jeffrey A. Robinson}

\begin{document}

\maketitle

\section{Introduction}

Go is a strongly typed programming language focused towards systems programming and concurrency.
We hope to support a subset of functionality of this language which include records, loops, conditionals, and functions.
We include a list of programs we want to support along with the grammar we plan on implementing.  

\section{Supported}

In our language we want to support enough operations so that we can construct a select number of algorithms.

\subsection{Expect to Support}

We want to make sure we can support any program an intro programming class covers.
To ensure this requirment we have a set of programs which must be supported which are located in the appendix.
We expect to be able to support the first three programs and hopefully the fourth.

\begin{itemize}
    \item [High Chance] Basic I/O
    \item [High Chance] Simple Calculator
    \item [High Chance] Iterative and Recursive Fibonacci
    \item [Low Chance] Infix to Prefix conversion
\end{itemize}

These programs require the following programming objects to be available to the programmer

\begin{itemize}
    \item primitives    - integers, reals, booleans, and strings
    \item conditionals  - if and else statements along with (possible) switch statements
    \item loops         - looping structure 
    \item functions     - non variadic functions with zero or more arguments and zero or one return values
    \item records       - arrays and structures
\end{itemize} 

\subsection{Might Support}

If we finish all required items and have enough time to implement the following 

\begin{itemize}
    \item full support of records   - Allow for grouping of data into single structure
    \item variadic functions        - Allow for functions allow a variadic number of arguments.
    \item multiple return values    - Allow for zero, one, or more return values and types
\end{itemize}
 
We believe that most of these are not difficult to implement but that due to time constraints they migth not be completed.

\subsection{Unlikely Support} 

In this section we talk about a list of items we want to support but will not have the time to support.

\begin{itemize}
    \item first class functions     - Allow for anonymous and named functions to created within a function.
    \item struct functions  - Functions for structures allow for implementation of interfaces.  However, we would need to support binding functions to structures and a method to call them.
    \item Interfaces        - Interfaces in Go use duck typing. This requires verification of structure function signatures  
    \item Imports           - Being able to handle file inclusion, dependency checks, and name collision. 
\end{itemize}

\section{Semantics}

Go supports static typing however it allows for untyped variables as in the snippet below

\begin{lstlisting}[title=Untyped value]
package main

func main() {
    i := 0
    println(i)
}
\end{lstlisting}

For this code snippet we have not defined the type of $i$ and we do not know if the value 0 is an $int$, $int32$, $int64$, or other numeric value.
If the value of the right side can be determined then the type of $i$ can be infered.
Otherwise, to ensure static typing we have the following rules

\begin{itemize}
    \item   If the expression on the right is an integral literal the type is $int$
    \item   If the expression on the right is a floating point literal the type is $float$
    \item   If the expression on the right is a string literal the type is $string$
    \item   If the expression on the right is an array of values then the type is an array of the least restrictive type as defined above. 
\end{itemize}

\appendix

\section{Appendix}

\subsection{Required Programs}

\subsection{Basic I/O}

\begin{lstlisting}[title=Basic I/O Program]
package main

func main() {
    print("Please enter a number")
    x := readFloat()
    print("Please enter another number")
    y := readFloat()
    println("The sum of these two numbers is", x + y);
}
\end{lstlisting}

\subsection{Basic Calculator}

\begin{lstlisting}[title=Basic Calculator]
package main

func add(x , y float64) float64 {
    return x + y
}

func subtract(x , y float64) float64 {
    return x - y
}

func multiply(x , y float64) float64 {
    return x * y
}

func divide(x , y float64) float64 {
    return x / y
}

func main() {
    println("Welcome to the calculator")
    for {
        

        err := false
        errMsg := ""
        x := readFloat()
        op := readCharacter()
        y := readFloat()
        readCharacter() // newline

        switch op {
            case '+':
                res = add(res,y)
            case '-':
                res = subtract(res,y)
            case '*':
                res = multiply(res,y)
            case '/':
                if y != 0 {
                    res = divide(res,y)
                } else {
                    err = true
                    errMsg = "Divide by zero"
                }
            default:
                err = true
                errMsg = "Invalid operator"
        }
        if err {
            println(errMsg)
            return
        } else {
            println(res)
        }
    }
}
\end{lstlisting}

\subsection{Iterative and Recursive Fibonacci}

\begin{lstlisting}[title=Iterative and Recursive Fibonacci]
package main

func recursiveFib(i int) int {
    if i < 2 {
        return 1
    }
    return fib(i-1)+fib(i-2)
}

func iterativeFib(i int) int {
    a := 1
    b := 1
    for i > 0 {
        t := a+b
        a = b
        b = t
        i--
    }
    return a
}

int main() {
    print(recursiveFib(10));
    print(iterativeFib(10));
}

\end{lstlisting}

\subsection{Infix to Prefix}

\begin{lstlisting}[title=Infix to Prefix]
package main

type stack struct {
    stk [100]rune
    pos   int
}

func (s *stack) Push(v rune) {
    s.stk[s.pos++] = v
}

func (s *stack) Pop() rune {
    s.pos--
    return s.stk[s.pos]
}

func (s *stack) Empty() bool {
    return s.pos == 0
}

func (s *stack) Peek() rune {
    return s.stk[s.pos - 1]
}

// Ruturns true if a has higher precedence
func precedence(a , b rune) bool {
    if a == b {
        return true
    }
    if a == '+' || a == '-' {
        if b == '*' || b == '/' {
            return false
        }
        return true
    }
    return !precedence(b,a)
}

func isDigit(r rune) bool {
    return '0' <= r && r <= '9'
} 

func reverse(data []rune) {
    left := 0
    right := len(data) - 1
    for left < right {
        t := data[left]
        data[left] = data[right]
        data[right = t
        left++
        right--
    }
}

int main() {
    data := {'1', '+', '2', '*' , '5', '/', '2' , '+' , '4' , '-' , '3' }
    result := []rune(nil)
    
    // Step 1 - Reverse array
    reverse(data)

    // Step 2 - Convert to postfix
    stk := &stack{}
    for i := range data {
        if isOp(data[i]) {
            for !stk.Empty() {
                if precedence(stk.Peek(),data[i]) {
                    result = append(result,stk.Pop())
                }else {
                    break
                }
            }else {
            stk.Push(data[i])
        } else {
            result = append(result,data[i])
        }
    }

    for !stk.Empty() {
        result = append(result,stk.Pop())
    }

    // Step 3 - reverse answer
    reverse(result)

    // Display answer
    for i := range result {
        print(result[i])
    }
    println()
}

\end{lstlisting}
\section{Grammar}

\input{./grammar/go--/go--.tex}


\end{document}